A Methodology and Implementation of Automated Emissions Harmonization for Use in Integrated Assessment Models

### 2018-01-30 paper.tex
finish up impl section

We herein present \code{aneris}' Python implementation and conceptual
workflow. The implementation is depicted in Figure \ref{fig:software}, and the
conceptual workflow is depicted in Figure \ref{fig:workflow}. The software is
composed of three primary components: the \texttt{HarmonizationDriver},
\texttt{Harmonizer}, and data processing routines (shown in red, green, and
blue, respectively in Figure \ref{fig:software-workflow}).

The \texttt{HarmonizationDriver} is an object designed to interface with
user-provided data and configuration files. Input data is assumed to be an Excel
file the in the in the IAMC format, i.e., using \code{Model}, \code{Scenario},
\code{Region}, \code{Variable}, and \code{Unit} columns in addition to columns
representing each modeled time period, which is the standard format within the
IAM community. It is responsible for downselecting data into unique combinations
for each model and scenario and then invoking the harmonization
process. Finally, it adds any exogenous emissions trajectories as required by
the user and recompiles the results and writes them to disc. The
\texttt{HarmonizationDriver} acts the primary interface for high-level users as
shown in Listing \ref{lst:tutorial}. Furthermore, a \gls{cli} is provided to
allow users to more easily incorporate the harmonization process in scripted
workflows (Listing \ref{lst:cli}, Figure \ref{fig:software}).

@@

The \texttt{Harmonizer} is a Python class whose concern is to harmonize various
model value trajectories given historical data and possible user method
\textit{overrides}, i.e., non-default methods. It is used by the
\texttt{HarmonizationDriver}; however it is also available to the user as a
first-class object. The \texttt{Harmonizer} requires that input data conform to
the \code{aneris} calculation data format, which expliciltly sepearates the
emissions species from the sector contributing the emissions (these are combined
in the single \code{Variable} column in the IAMC format). Because the
\texttt{Harmonizer} is configure for a single instance of a model and scenario,
the canonical data format includes columns named \texttt{region},
\texttt{sector}, \texttt{gas}, and \texttt{units}. Once configured with
appropriate input data (model and history) as well as potential method
overrides, the \texttt{Harmonizer}'s \texttt{harmonize()} method can be invoked
which returns a \texttt{pandas.DataFrame} of harmonized data. The object can
additionally be queried directly as to its \texttt{default\_methods()},
\texttt{methods()} (i.e., methods used with overrides), and \texttt{metadata()}
(i.e., methods used with all branching information along each path in the
decision tree).

There are also a variety of tools and utilities provided to users and also used
by the \texttt{HarmonizationDriver} in order to process both input and output
data. These include an \texttt{EmissionsAggregator} class and related routines
used to combine historical emissions to native model regions (where historical
data is defined at a higher spatial resolution than a model; see, e.g., Figure
\ref{fig:regions}) and a \texttt{FormatTranslator} class which defines an
interface for translating \texttt{pandas.DataFrame}s between the IAMC format
expected for input and output data and the calculation format used by
\code{aneris}' \texttt{Harmonizer}.

\begin{figure}
  \subfloat[Python Implementation\label{fig:software}]{
    \includegraphics[width=0.775\textwidth]{implementation_schematic.pdf}
    }
  \subfloat[Workflow\label{fig:workflow}]{
    \includegraphics[width=0.225\textwidth]{aneris-workflow.pdf}
    }
    \caption{ Panel \ref{fig:software} shows the various objects in the
      \code{aneris} codebase as well as a short description of their scope of
      concern. Panel \ref{fig:workflow} displays the full harmonization process
      as executed by \code{aneris} for a single instance of a model and
      scenario. Operations that can be configured with user-based input
      configurations are shown in purple. Operations goverened by the
      \texttt{HarmonizationDriver} are shown in red. Data processing operations
      are shown in blue. The core harmonization process, governed by the
      \texttt{Harmonizer} is shown in green.  }
    \label{fig:software-workflow}
\end{figure}

Users are able to control the harmonization process via a number of options
(with examples provided
\href{http://mattgidden.com/aneris/config.html}{online}). The primary mechanism
by which users control the process is by providing \textit{override} methods for
any combination of region and variable (i.e., sector and gas species).  In
practice, it may be possible that not all default methods chosen will provide
robust harmonized trajectories, especially if there is a significant difference
between historical and model values in the harmonization year, if there is
significant upward or downward movement in the model trajectory, or if there are
known discrepancies in sectoral definition between the IAM and historical data
source. In such cases, override methods can ameliorate the issues associated
with the default method choice.

In order to help identify these cases, harmonization \textit{diagnostics} are
provided which analyze the relative difference between harmonized and
unharmonized trajectories at their mid and end-points. These values are
configurable by the user, but defaults of $400$\% and $200$\%, respectively, are
provided based on experiences of using \code{aneris} to date. If override
methods are provided, they are used instead of the default methods as determined
by the default method decision tree. Furthermore, users can set the
above-mentioned thresholds as well as the LUC method used in the decision tree
(see Figure \ref{fig:decision_tree}).

Input data then undergoes a cleaning operation, which adds (null) model
trajectories that exist in the historical data set but are not provided by the
model input and detects any issues that would cause the harmonization process to
fail. The methods used to harmonize the data are then determined and the
harmonization process is executed. Upon completion of the harmonization process,
aggregation of common analysis regions is performed. A common regional
aggregation used in the IAM community was defined in the Representative
Concentration Pathways (RCPs) \cite{vuuren_representative_2011}, shown in Figure
\ref{fig:regions}. Finally, any exogenous trajectories the user provides are
added. Exogenous trajectories are normally provided for unmodeled gases with
well-accepted scenario trajectories, e.g., chlorofluorocarbons provided by WMO
\cite{wmo2014}. Upon completion, the harmonized trajectories and meta data
regarding the harmonization process are returned. A description of all returned
meta data is provided in Table \ref{tab:metadata}.

@@

\begin{table}[]
\centering
\caption{Meta data provided by the \code{aneris} harmonization routine. This meta data is provided for every combination of region, sector, and emissions species.}
\label{tab:metadata}
\begin{tabular}{|p{2cm}|p{8cm}|}
\hline
\textbf{Column}       & \textbf{Description}                                      \\
\hline
\hline
method       & The harmonization method used.                                               \\
\hline
default      & The default harmonization method as determined by the default decision tree. \\
\hline
override     & The method provided as an override (if any).                                 \\
\hline
offset       & The offset value between history and model in the harmonization year.        \\
\hline
ratio        & The ratio value between history and model in the harmonization year.         \\
\hline
cov          & The coefficient of variation value of the historical trajectory.                           \\
\hline
unharmonized & The unharmonized value in the harmonization year.                            \\
\hline
history      & The historical value in the harmonization year.                             \\
\hline
harmonized   & The resulting harmonized value in the harmonization year.\\
\hline
\end{tabular}
\end{table}
\section{Results}\label{sec:results}

>>>>

We herein present \code{aneris}' Python implementation, depicted in Figure
\ref{fig:software}, and conceptual, depicted in Figure \ref{fig:workflow}. The
library is composed of a number of utilities as well as three primary
components: the \texttt{HarmonizationDriver}, \texttt{Harmonizer}, and data
processing routines (shown in red, green, and blue, respectively in Figure
\ref{fig:software-workflow}).

The \texttt{HarmonizationDriver} is an object designed to interface with
user-provided data and configuration files. Input data (i.e., unharmonized model
results) is assumed to be an Excel file in the standard data format within the
IAM community, i.e., with \code{Model}, \code{Scenario}, \code{Region},
\code{Variable}, and \code{Unit} columns in addition to columns representing
each modeled time period. It is responsible for down-selecting data into
separate datasets for each model and scenario, invoking the harmonization
process on each dataset, and recompiling the results. The
\texttt{HarmonizationDriver} acts the primary interface for high-level users as
shown by the usage of the \texttt{driver} object in Listing
\ref{lst:tutorial}. Furthermore, a \gls{cli} is provided to allow users to more
easily incorporate the harmonization process in scripted workflows (Listing
\ref{lst:cli}, Figure \ref{fig:software}).

@@

The \texttt{Harmonizer} is a Python class whose concern is to harmonize model
value trajectories given historical data and possible user method
\textit{overrides}, i.e., non-default methods (described further below). It is
used by the \texttt{HarmonizationDriver}; however it is also available to the
user as a first-class object. The \texttt{Harmonizer} requires that input data
conform to the \code{aneris} calculation data format, which explicitly separates
the emissions species from the sector contributing the emissions (these are
combined in the single \code{Variable} column in the standard IAM
format). Because the \texttt{Harmonizer} is designed to operate on a single
instance of a model and scenario, the canonical data format includes
\texttt{region}, \texttt{sector}, \texttt{gas}, and \texttt{units} columns
without extraneous metadata columns for the model and scenario. Once configured
with appropriate input data (model and history) as well as potential method
overrides, the \texttt{Harmonizer}'s \texttt{harmonize()} method can be invoked
which returns a \texttt{pandas.DataFrame} of harmonized data. The object can
additionally be queried directly as to its \texttt{default\_methods()},
\texttt{methods()} (i.e., methods used with overrides), and \texttt{metadata()}
(i.e., methods used with all branching information along each path in the
decision tree).

There are also a variety of tools and utilities provided to users and also used
by the \texttt{HarmonizationDriver} in order to process both input and output
data. These include an \texttt{EmissionsAggregator} class and related routines
used to generate sectoral emissions totals, generate regional totals, and
combine historical emissions to native model regions (where historical data is
defined at a higher spatial resolution than a model; see, e.g., Figure
\ref{fig:regions}). A \texttt{FormatTranslator} class is also provided which
defines an interface for translating \texttt{pandas.DataFrame}s between the IAM
format expected for input and output data and the calculation format used by
\code{aneris}' \texttt{Harmonizer}.

\begin{figure}
  \subfloat[Python Implementation\label{fig:software}]{
    \includegraphics[width=0.775\textwidth]{implementation_schematic.pdf}
    }
  \subfloat[Workflow\label{fig:workflow}]{
    \includegraphics[width=0.225\textwidth]{aneris-workflow.pdf}
    }
    \caption{ Panel \ref{fig:software} shows the various objects in the
      \code{aneris} code base as well as a short description of their scope of
      concern. Panel \ref{fig:workflow} displays the full harmonization process
      as executed by \code{aneris} for a single instance of a model and
      scenario. Operations that can be configured with user-based input
      configurations are shown in purple. Operations governed by the
      \texttt{HarmonizationDriver} are shown in red. Data processing operations
      are shown in blue. The core harmonization process, governed by the
      \texttt{Harmonizer} is shown in green.  }
    \label{fig:software-workflow}
\end{figure}

The full harmonization workflow, outlined in Figure \ref{fig:workflow}, begins by
cleaning input data, which adds (null) model trajectories that exist in the
historical data set but are not provided by the model input and detects any
issues that would cause the harmonization process to fail. The methods used to
harmonize the data are then determined and the harmonization process is
executed. Upon completion of the harmonization process, aggregation of common
analysis regions is performed. A common regional aggregation used in the IAM
community was defined in the Representative Concentration Pathways (RCPs)
\cite{vuuren_representative_2011}, shown in Figure \ref{fig:regions}. Finally,
any exogenous trajectories the user provides are added. Exogenous trajectories
are normally provided for unmodeled gases with well-accepted scenario
trajectories, e.g., chlorofluorocarbons provided by WMO \cite{wmo2014}. Upon
completion, the harmonized trajectories and meta data regarding the
harmonization process are returned. A description of all returned meta data is
provided in Table \ref{tab:metadata}.

@@

\begin{table}[]
\centering
\caption{Meta data provided by the \code{aneris} harmonization routine. This meta data is provided for every combination of region, sector, and emissions species.}
\label{tab:metadata}
\begin{tabular}{|p{2cm}|p{8cm}|}
\hline
\textbf{Column}       & \textbf{Description}                                      \\
\hline
\hline
method       & The harmonization method used.                                               \\
\hline
default      & The default harmonization method as determined by the default decision tree. \\
\hline
override     & The method provided as an override (if any).                                 \\
\hline
offset       & The offset value between history and model in the harmonization year.        \\
\hline
ratio        & The ratio value between history and model in the harmonization year.         \\
\hline
cov          & The coefficient of variation value of the historical trajectory.                           \\
\hline
unharmonized & The unharmonized value in the harmonization year.                            \\
\hline
history      & The historical value in the harmonization year.                             \\
\hline
harmonized   & The resulting harmonized value in the harmonization year.\\
\hline
\end{tabular}
\end{table}

Users are able to control the harmonization process via a number of options
(with examples provided
\href{http://mattgidden.com/aneris/config.html}{online}). The primary mechanism
by which users control the process is by providing \textit{override} methods for
any combination of region and variable (i.e., sector and gas species).  In
practice, it may be possible that not all default methods chosen will provide
robust harmonized trajectories, especially if there is a significant difference
between historical and model values in the harmonization year, if there is
significant upward or downward movement in the model trajectory, or if there are
known discrepancies in sectoral definition between the IAM and historical data
source. In such cases, users can override default methods with their method of
choice and both the default method and override is reported in the resulting
metadata.

In order to help identify cases where overrides may be of interest,
harmonization \textit{diagnostics} are provided which analyze the relative
difference between harmonized and unharmonized trajectories at their mid and
end-points. These values are configurable by the user, but defaults of $400$\%
and $200$\%, respectively, are provided based on experiences of the authors'
using \code{aneris} to date.

\section{Results}\label{sec:results}
